Microxanox: an R package for simulating an aquatic \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.

### 2022-05-10 inst/manuscript/manuscript_microxanox.tex
Additional references in bibliograhy and beautified graphs.

Many ecosystems are exposed to gradual changes of environmental
variables, to which the responses are not always as gradual as the
change of the environmental variable \citep{REF_NEEDED}. One example
where the gradual change of a single environmental variable is causing
an abrupt change of the system is the switch from an aerobic to
anaerobic system. This system has been investigated by \citet{Bush2017}.
We wanted to take this investigation one step further, and look at the
role an increased biodiversity plays in these dynamics
\citep{REF_NEEDED}. For this purpose, we developed this package.

The microxanox package is a package for simulating a three functional
group system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur
bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical
substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced
sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
biogeochemical processes and is based on \citep{Bush2017} (See
\citep{Bush2017} for a detailed discussion of the model).

The aims of the \texttt{microxanox} package are twofold: firstly, to
reproduce the results shown by \citep{Bush2017}, which is accomplished
in the \href{LINK\%20NEEDED}{vignette Partial reproduction of Bush et
al} and secondly, to take these results one step further, to includes
new functionality to address our research question as published in
\citep{REF_NEEDED}.

For this, we extended the model and added functionality for:

\begin{itemize}
\tightlist
\item
  Multiple strains (effectively unlimited) per functional group.
\item
  Adding temporally varying oxygen diffusivity.
\item
  Adding random noise in substrate concentrations.
\item
  Including immigration.
\item
  Setting minimum population abundances.
\end{itemize}

In addition to the model itself, the package includes some functions to
analyse the results as well as visualize the results to provide a
starting point for customized visualizations based on own requirements.

@@

We will discuss the general structure and functionality of the package
here without going into to much detail. A more detailed discussion can
be found in the \href{LINK_NEEDED}{User Guide}.

The model by \citet{Bush2017} is based on ODEs for the rates of change
which are specified in the function \texttt{bushplus\_dynamic\_model()}.
This augmented version of the model published in \citep{Bush2017} can
handle multiple strains within each of the three functional groups,
temporal variation in oxygen diffusivity, and events.

The package is not intended to provide a modelling framework which can
be adjusted easily to all needs, but primarily a tool to implement the
model used by \citep{Bush2017} and to extend it to our needs
\citep{REF_NEEDED}. Consequently, any more substantial changes and
adaptations are likely to need a change in the source code.

Nevertheless, the source code is structured in modular, so that e.g.~the
event definition can easily be changed, or other aspects can be
adjusted. All values in the parameter object can be changed as needed
and the general structure of the code should make it not to difficult to
adapt the model to other similar systems.

\hypertarget{software-architecture}{%
\subsection{Software Architecture}\label{software-architecture}}

\begin{quote}
\{\textgreater\textgreater Give a short overview of the overall software
architecture; provide a pictorial component overview or similar (if
possible). If necessary provide im- plementation
details.\textless\textless\}
\end{quote}

The framework used when writing this package aims at reproducibility of
the results (see \citep{REF_NEEDED} for examples). It builds on the
following main considerations:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  all parameter needed to run a simulation or find a stable state are
  contained in a single parameter object. This object is created by
  using the functions \texttt{new\_...\_parameter()},
  \texttt{new\_initial\_state()} and \texttt{new\_strain\_parameter()}.
  Which one of the \texttt{new\_...\_parameter()} functions has to be
  used when, will be discussed in the section @ref(runsim) and in more
  detail in the \href{LINK_NEEDED}{User Guide}.
\item
  The function call \texttt{run\_...(parameter)} will run the simulation
  using the parameter as defined in the object \texttt{parameter}.
\item
  The return value of the \texttt{run\_...(parameter)} function is
  identical to the parameter object plus an additional slot named
  \texttt{results} which contains the results of the run
\item
  As the return value contains all parameter, it is possible to re-run
  the simulation by simply running \texttt{run\_...(result)}.
\end{enumerate}

The point that the results object contains all parameter needed to run
the simulation, promotes reproducibility and makes incremental changes
of individual parameters and re-running the simulations much easier, as
can be seen in our research and the RMarkdown \citep{REF_NEEDED} file
used to run and analyse the experiments \citep{REF_NEEDED}.

In the following sections we describe the general usage of the package:
running one simulation, finding steady states across an environmental
gradient, calculating measures of stability, and visualization.

\hypertarget{running-one-simulation}{%
\subsubsection{Running one simulation}\label{running-one-simulation}}

@@

The individual simulation (\texttt{run\_simulation()} function) is the
working horse in this package. In this function, the ODEs are solved
using the function \texttt{ode} in the package deSolve package
\citep{Soetaert2010}. The \texttt{run\_simulation()} function needs only
one argument - an object as created by the function
\texttt{new\_runsim\_parameter()}. The parameter object contains among
other parameter the \texttt{strain\_parameter} which can be created by
the function \texttt{new\_strain\_parameter()}. For a detailed
description of the parameter, their meaning and how they are created and
can be set and changed please see the User Guide which accompanies the
package or is available at \href{@LINK_NEEDED}{User Guide}

@@

@@

\hypertarget{finding-a-steady-state-of-the-model}{%
\subsubsection{Finding a Steady State of the
model}\label{finding-a-steady-state-of-the-model}}

\textcolor{red}{@Owen: can you please fill in the references to these mathods? Thanks.}

There are two methods for finding steady states implemented in this
package. The first runs a separate simulation for each combination of
starting conditions and oxygen diffusivity (let us term this the
\emph{Replication method})
\textbackslash textcolor\{red\}\{\{\textgreater\textgreater{}\citet{Owen}
- does this approach has a name, and do you have a reference where it
has been used or described?\textless\textless\}. The second runs only
two simulations, with step-wise and slowly temporally increasing or
decreasing oxygen diffusivities and recorded of state just before change
to a new oxygen diffusivity (let us term this the \emph{Temporal
method})\textbackslash textcolor\{red\}\{\{\textgreater\textgreater{}\citet{Owen}
- Has this been used in other publications? I do not have a clue what I
could search for to find out\ldots\textless\textless\}.

\hypertarget{replication-method}{%
\paragraph{Replication Method}\label{replication-method}}

\hypertarget{temporal-method}{%
\paragraph{Temporal Method}\label{temporal-method}}

The temporal method involves two simulations for a particular system
configuration (parameter set). In one simulation the oxygen diffusivity
is \emph{increased} in a step-wise fashion. In the other it is
\emph{decreased} in a step-wise fashion. That is, oxygen diffusivity is
held at a constant level for long enough for steady state to be reach,
that state is recorded, and then a slightly higher (or lower) oxygen
diffusivity value is set. Hence, at that time point, the system is
effectively started with initial conditions that are the state of the
system in the previous time step
\textcolor{red}{{>>@Owen: is there a ref for this approach?<<}}.

This is implemented in the function \texttt{run\_temporal\_ssfind()},
which takes a parameter object as created by the function
\texttt{new\_temporal\_ssfind\_parameter()}. It is planned for a later
release, to run these two simulations in parallel.

\hypertarget{extracting-stability-measures}{%
\subsubsection{Extracting Stability
Measures}\label{extracting-stability-measures}}

From the raw results returned by these \texttt{run\_...()} functions,
the stability measures can be extracted by using the function
\texttt{get\_stability\_measures()}. These measures include
non-linearity and hysteresis measures, of the response of the simulated
system to environmental change
\textcolor{red}{{>>@Owen: is there a paper, so that we can say "as e.g. used in REF?<<}}.

\hypertarget{use-cases}{%
\subsection{Use Cases}\label{use-cases}}

We will now show three use cases which illustrate applications of the
package. All use cases can be used as starting points for other
investigations of the behavior of the system.

The first two use cases are described in detail in the User Guide and
the Partial Reproduction Vignettes. The third is taken from
\citet{REF_NEEDED} for which this R package was designed. All of these
use cases can be expanded to larger numbers of strains per functional
group and variable values van be changed.

\hypertarget{regime-shifts-during-temporal-environmental-change}{%
\subsubsection{Regime shifts during temporal environmental
change}\label{regime-shifts-during-temporal-environmental-change}}

\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}

In the \href{LINK_NEEDED}{User Guide} we used a one strain system
(section ``1 strain per functional group'') and three strain system
(section ``3 strains per functional group'') to determine as an example
the stable states during temporal environmental changes (the oxygen
diffusivity). From these simulatiuons, we extracted measures of
nonlinearity and hysteresis. See Fig @ref(fig:plot-dynamics) as an
example plot of the simulations.

{\centering \includegraphics[width=350px]{figures/ug_three_strains_stable_state}

\caption{Plot of the stable states of the simulation runs under different oxygen diffusivity. The top graph are the Organisms (ech initially with three strains) while the lower graph is the substrate availability under the same oxygen diffusivities. Details can be found in the "User Guide" section "Three strains per functional group".}\label{fig:uc1_stable_state}
\end{figure}

We saved all parameter and result objects, so that all results are
completely reproducible. The general flow of the experiment is identical
to the one shown in @ref(fig:runsim\_example).

\hypertarget{the-extent-of-hysteresis-depends-on-community-composition}{%
\subsubsection{The extent of hysteresis depends on community
composition}\label{the-extent-of-hysteresis-depends-on-community-composition}}

\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}

One of the reasons to develop this package was to reproduce the results
presented in \citet{Bush2017}. This was achieved as demonstrated in the
\href{LINK_NEEDED}{Partial Reproduction supplement}. All aspects in the
paper could be reproduced and are shown in the vignette.

{\centering \includegraphics[width=350px]{figures/user_guide_hysteresis}

\caption{Hysteresis of all assessed combinations of variability.}\label{fig:user_guide_hysteresis}
\end{figure}

As all objects were saved, results can be easily reproduced and
parameter can be changed to assess the impact of these changes (e.g.~to
conduct a sensitivity analysis).

\hypertarget{effects-of-functional-diversity-on-regime-shifts}{%
\subsubsection{Effects of functional diversity on regime
shifts}\label{effects-of-functional-diversity-on-regime-shifts}}

\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}

As discussed in the paper \citep{REF_NEEDED}, the role biodiversity
plays in abrupt regime shifts based on gradual changing environmental
parameter is not well understood. This model (as part of the package)
has been used to investigate these dynamics and the results are
available in with the DOI DOI NEEDED \citet{REF_NEEDED}.

\textcolor{red}{{>>@Owen: More details from the paper. I will approach Romana when the first draft is ready so that she can put something in, onless you will add a few sentences here.<<}}

@@

The free and open source implementation and extension of the model used
in \citet{Bush2017} provides the means of reproducing the results
published while at the same time provides the means of doing
investigations which extend the original publication, as done in
\citet{REF_NEEDED}.

The combination of all parameter in a single parameter object as well as
the return of the simulation as a result object which inherits from the
parameter object and adds a results slot provides a relatively easy to
use framework to implement reproducible experiments.

\textcolor{red}{{>>@Owen: Could you put a few sentence in highlighting the impact this package could have considering the importance of possible research questions?<<}}

@@

This package makes an important paper
\textcolor{red}{{>>@Owen: is this correct?<<}} \citep{Bush2017} fully
reproducible without having to re-implement the code. By doing so, it
opens the door for further investigations based on the system as done in
\citet{REF_NEEDED} and, consequently, increases the reach (and ultimaly)
the impact of the original paper by \citet{Bush2017}. This demonstrates
again \citep{REF_NEEDED} the importance of publishing not only the data
but also the code used in a publication.

\textcolor{red}{{>>@Owen: could you write a short conclusion paragraph about the ecological relevance of `microxanox`?<<}}

In addition to the ecological relevance and importance of this package
as discussed above, it can be used as a template for the implementation
of other different models. By doing so, other models can profit from the
reproducibility aspects as well as the flexibility as implemented in
this package.

@@

We wish to confirm that there are no known conflicts of interest
associated with this publication and there has been no significant
financial support for this work that could have influenced its outcome.

@@

Optionally thank people and institutes you need to acknowledge.

>>>>

Mathematical models play a key role in the development of understanding
about how ecosystems work and how they respond to environmental changes
\citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also
critical for developing hypotheses that can then be tested in
experimental studies. One area of ecology in which simple models have
played a very influential role is the area of how ecosystems respond to
gradual environmental change \citep{REF_NEEDED}. - Scheffer, M.,
Carpenter, S., Foley, J. a, Folke, C. \& Walker, B. (2001). Catastrophic
shifts in ecosystems. Nature, 413, 591--6.

It is conceivable that an ecosystem state, such as the total biomass of
a particular type of bacteria, may remain unchanged when something about
its environment changes, such as the temperature. It is possible that
the ecosystem state changes gradually itself. It is also possible that
the ecosystem state may change abruptly into a new state that is
difficult to recover from \citep{Scheffer2001}.

This possibility for abrupt, perhaps catastrophic changes that are
difficult to reverse causes considerable concern
\citep{Collins2021, Northrop2021, Vandermeer2019}.

An example where the gradual change of a single environmental variable
is causing an abrupt change of the system is the switch from an aerobic
(oxygen is available for metabolism) to anaerobic (oxygen generally
unavailable) system. This system has been investigated by
\citet{Bush2017} in a simulation study of a mathematical model of the
system. They showed that gradual change in the environment could cause
catastrophic changes in the ecosystem state that would be difficult to
reverse.

One limitation of the study by \citet{Bush2017} was limited biodiversity
in the ecosystem model. Specifically, there were three functional groups
of bacteria, but within each of these groups, there was no biodiversity.
This then leaves open the question of if and how biodiversity within
functional groups, in their model, affects the ecosystem response to
environmental change. This limitation is not limited to the study of
\citet{Bush2017}, there are few if any studies of the effects of
biodiversity on abrupt transitions between ecosystem states.

We decided to fill this research gap, by making a simulation study of
how within functional group biodiversity affects ecosystem responses to
environmental change \citet{Limberger2022}, and to base our work on the
work and model of \citet{Bush2017}. It was with this goal in mind that
we developed the \emph{microxanox} package. The first stage of
development was to write code from scratch (as there was no available
code to start from) and to confirm that this new implementation would
reproduce the previously published results. The resulting reproduction
is available as one of the package vignettes:
\href{LINK\%20NEEDED}{vignette Partial reproduction of Bush et al}.

The second stage was to add functionality that would be necessary to
answer our research question. Most importantly, we made it possible to
have multiple species of bacteria within each of the three functional
groups, for the multiple species to differ in their characteristics, and
to vary the number of species and amount of variability among them. We
also added functionality that allowed: temporally varying environmental
conditions, addition of random noise to state variables, and
immigration. In addition to the model itself, the package includes some
functions to analyse the results as well as visualize the results to
provide a starting point for customized visualizations based on own
requirements. The basic and additional functionality is described in the
package \href{LINK_NEEDED}{User Guide}.

@@

The \emph{microxanox} package is for simulating a three functional group
system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur
bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical
substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced
sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
biogeochemical processes and is based on \citet{Bush2017} (See
\citet{Bush2017} for a detailed discussion of the model). At the core of
the simulations is a set of ordinary differential equations (specified
in the function \texttt{bushplus\_dynamic\_model()}, though users do not
need to call this function directly). There are functions for running
individual simulations and for running a set of simulations across, for
example, a range of environmental conditions.

To make the simulation run with multiple species, we expressed different
species characteristics in the elements of vectors and matrices. We also
adapted the code for the ordinary differential equations to include the
vectors and matrices, and to use matrix mathematics. In this way, we
made it possible to run simulations with different numbers of species
without having to change the underlying code.

The package functions and code are structured in modular, so that some
new functionality can be easily added. E.g. the event definition can
easily be changed, or other aspects can be adjusted. All values in the
parameter object can be changed as needed and the general structure of
the code should make it not to difficult to adapt the model to other
similar systems.

\hypertarget{software-architecture}{%
\subsection{Software architecture}\label{software-architecture}}

\begin{quote}
\{\textgreater\textgreater Give a short overview of the overall software
architecture; provide a pictorial component overview or similar (if
possible). If necessary provide implementation
details.\textless\textless\}
\end{quote}

The framework used when writing this package aimed to maximise
simplicity for the user, and to make it easy to reproduce results (see
the supplement to \citet{Limberger2022} for an example of how this is
used). As such, all the parameters needed to run a simulation or find a
stable state are contained in a single object (which can easily be
created using included functions). This parameter object is given to a
function that runs the simulations and returns the results. The returned
results object is identical to the parameter object but with an
additional slot named \texttt{results}, which contains the results of
the run. Thus the returned results object contains the simulation
conditions (parameters) as well as the rsults, and can be used to run
the simulation again. This promotes reproducibility and makes
incremental changes of individual parameters and re-running the
simulations straightforward.

In the following sections we describe how to use the package to run one
simulation and to find steady states across an environmental gradient.

\hypertarget{running-one-simulation}{%
\subsection{Running one simulation}\label{running-one-simulation}}

@@

A simulation is run using the \texttt{run\_simulation()} function. In
this function, the ODEs are solved using the function \texttt{ode()} in
the package \emph{deSolve} package \citep{Soetaert2010}. The
\texttt{run\_simulation()} function needs only one argument - an object
as created by the function \texttt{new\_runsim\_parameter()}. The
parameter object returned by \texttt{new\_runsim\_parameter()} contains
among other things the \texttt{strain\_parameter} object, which can be
created by the function \texttt{new\_strain\_parameter()}. For a
detailed description of the parameter objects, their meaning and how
they are created and have values set and changed please see the
\emph{User Guide} which accompanies the package or is available at
\href{@LINK_NEEDED}{User Guide}

@@

\hypertarget{finding-steady-states}{%
\subsection{Finding Steady States}\label{finding-steady-states}}

The general approach used to find the stable state of the system with a
specific parameter set is to run the simulation for a long time and
record the final state. When one does this across a range of
environmental conditions, one discovers how the steady state of the
system responds to the environmental conditions. The package contains
functionality for finding steady states that correspond to values of one
environmental variable, namely the value of oxygen diffusivity.

Two methods for finding steady states are implemented. The first runs a
separate simulation for each combination of starting conditions and
oxygen diffusivity (let us term this the \emph{Replication method}).
This is the method used in the \citet{Bush2017} study. The second runs
only two simulations, with step-wise and slowly temporally increasing
oxygen diffusivity, and the other with step-wise and slowly decreasing
or decreasing oxygen diffusivity. During this temporal environmental
change, the state of the system was recorded just before change to a new
oxygen diffusivity (let us term this the \emph{Temporal method}). We
implemented two methods since there is no definitive best method, and in
order to check if results were sensitive to choice of method.

This temporal method implemented in the function
\texttt{run\_temporal\_ssfind()}, which takes a parameter object as
created by the function \texttt{new\_temporal\_ssfind\_parameter()}. It
is planned for a later release, to run these two simulations in
parallel.

\hypertarget{visualising-and-analysing-results}{%
\subsection{Visualising and analysing
results}\label{visualising-and-analysing-results}}

From the raw results returned, summary measures about how the ecoystem
stable states respond to environmental change can be extracted. The
function \texttt{get\_stability\_measures()} returns quantities such as
the amount of environmental change required for the system to abruptly
change to a different state.

@@

@@

The free and open source implementation and extension of the model used
in \citet{Bush2017} provides the means of reproducing the results
published while at the same time provides the means of doing unique,
innovative, and important investigations of how ecosystems respond to
environmental change. The design of the package code and functionality
is with reproducibility in mind: the combination of all parameters being
in a single parameter object as well as the return of the simulation as
a result object which inherits from the parameter object provides a
relatively easy to use framework to implement reproducible experiments.

While the package was not intended to provide a code which can be very
easily adapted to other types of ecosystems and types of organism, it
seems likely that the overall framework could be used as a basis for
models of other ecosystems and organisms. For example, all such models
would have parameters that differ among species and need to be described
in an object. All would need to run simulations and sets of simulations
across environmental conditions. Researchers want to model a new
ecosystem do not, therefore, have to start from scratch. Nevertheless,
depending on how different were the new ecosystem to model, the require
code changes could be quite significant.

We evidence the impact of the \emph{microxanox} package by describing
three use cases, each of which was only possible with the package and
that contains new results. The first two use cases are described in
detail in the User Guide and the Partial Reproduction Vignettes. The
third is taken from \citet{REF_NEEDED} for which this R package was
designed. All of these use cases can be expanded to larger numbers of
strains per functional group and variable values van be changed.

\hypertarget{use-case-1-regime-shifts-during-temporal-environmental-change}{%
\subsection{Use case 1: Regime shifts during temporal environmental
change}\label{use-case-1-regime-shifts-during-temporal-environmental-change}}

The package contains functionality to make a specific pattern of
temporal change in the environmental condition oxygen diffusivity; this
functionality forms the basis of the temporal method for finding stable
states. A simple example of this functionality is given in the Partial
Reproduction vignette, which we briefly show here (Figure @ref(fig:uc1).
The example is composed of a single simulation, at the beginning of
which the system is in the oxic state with high abundance of
cyanobacteria. Oyxgen diffusivity is then slowly decreased and
eventually, around hour 30'000 the system switches to the anoxic state,
with high abundance of both sulfur bacteria types. The oxygen
diffusivity is then quickly increased and at around hour 38'000 the
system abruptly switches back to the oxic state.

{\centering \includegraphics[width=350px]{./figures/gen_uc1_partrep_temporal_state_switching}

\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes. Plot of the stable states of the simulation runs under different oxygen diffusivity.}\label{fig:uc1}
\end{figure}

\hypertarget{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}{%
\subsection{Use case 2: The extent of hysteresis depends on community
composition}\label{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}}

The package also contains a function to extract summary features of
ecosystem responses to environmental change, such as the amount of
hysteresis displayed by the ecosystem. Hysteresis is a key feature of
ecosystem responses to environmental change, because it is related to
how difficult it can be to reserve the effects of environmental change
\citep{Scheffer2001}. The amount of hysteresis was measured here as the
extent of the environmental condition for which there were two stable
states. I.e. it was the extent of the environmental conditions for which
historical conditions play an important role in determining the current
system state (a definition of hysteresis). The results show that the
amount of hysteresis depends greatly on the combinations of organisms
present (Figure @ref(fig:uc2)). For example, with only the CB
(cyanobacteria) present, there was no hysteresis. In contrast, the
presence of both CB and SB (sulfate reducing bacteria) led to a large
amount of hysteresis. (These results are also given in the Partial
Reproduction vignette.)

\textbf{\citet{Rainer}\ldots{} the figure needs beautifying please.}
\textbf{\citet{Owen}: is this OK like this?}

{\centering \includegraphics[width=350px]{./figures/gen_uc2_user_guide_hysteresis}

\caption{The amount of hysteresis depends on the combination of types of organisms present.}\label{fig:uc2}
\end{figure}

\hypertarget{use-case-3-effects-of-functional-diversity-on-regime-shifts}{%
\subsection{Use case 3: Effects of functional diversity on regime
shifts}\label{use-case-3-effects-of-functional-diversity-on-regime-shifts}}

As discussed in the Introduction section, the package was largely
motivated by the question of how biodiversity influences ecosystem
responses to environmental change. Extensive results concerning this
question are given in a separate publication \citet{Limberger2022}. Here
we describe one of the results, which is that having biodiversity in a
functional group can allow state changes to occur that otherwise would
not have. I.e. biodiversity can qualitatively change the state of the
ecosystem.

Biodiversity was added to the functional groups using the
\texttt{new\_strain\_parameter()} function to create a parameter set
with multiple species per functional group (albeit all with identical
features) and then to add variability among the species by calling the
\texttt{add\_strain\_var()} function. This function takes an already
existing parameter set and adds the specified about of variation. The
new parameter object is then used as before.

Figure @ref(fig:uc3) shows a simulation with two species (strains) in
each of the three functional groups. The ecosystem starts in the oxic
state, though with relatively high abundance of each functional group.
The strain of SB that is more tolerant to oxygen (SB\_1) initially
decreases in abundance, but then increases, and the other (SB\_2) strain
then becomes abundance and SB\_1 declines. Furthermore, the
cyanobacteria crash in abundance, and the system switches to the anoxic
state. Importantly, the switch does not occur if there are two identical
strains with tolerance half way between those in Figure @ref(fig:uc3).

\textbf{Rainer\ldots{} please beautify the figure as required. It comes
from the Limberger et al supplement, Figure 5.4} \textbf{\citet{Owen}:
is this OK like this? Used the graph from the supplement}

{\centering \includegraphics[width=350px]{figures/uc3_supplement_5_4}

\caption{The dynamics of the ecosystem when there are two species in each functional group, and some variation (diversity) in species parameters.}\label{fig:uc3}
\end{figure}

@@

\emph{microxanox} allows the simulation, visualisation, and analysis of
a model of a microbial ecosystem while allowing variation in the amount
of diversity containing in each of the functional groups of organisms
present. It has been used for the research described in another paper of
ours that is one of the first investigations of the effects of diversity
on ecosystem resilience \citet{Limberger2022}. In that paper, we show
that diversity can have large and important effects of ecosystem
responses, highlighting the need for models such as ours, within which
one can easily manipulate the amount of biodiversity. The
\emph{microxanox} package has also been used to reproduce the results of
the paper that inspired the package development \citet{Bush2017}.

The package greatly lowers the amount of work required in further
investigations of the specific ecosystem modelled. There has, for
example, been quite limited investigation of how biodiversity influences
the short-term responses of the modelled ecosystem to environmental
change. Likewise, the package could be used to power an investigation of
the effects of biodiversity on the usefulness of early warning signals
of abrupt ecosystem change \citep{Scheffer2009}. In addition this
package could be used as a template for the implementation for
developing models of other types of ecosystems and organism. By doing
so, other models can profit from the overall framework used, and the
reproducibility aspects as well as the flexibility implemented.

@@

The authors declare no known conflicting or competing interests
associated with this publication and there has been no significant
financial support for this work that could have influenced its outcome.

@@

This project was part of SNF Project 310030\_188431. The project was
also supported by the University of Zurich Research Priority Programme
in Global Change and Biodiversity.
